---
title: "PS 15 Section 8 - Multivariate regression (part 2)"
output: html_document
---

## REMINDERS
1. KNIT AS YOU GO!
2. ALWAYS SET YOUR WORKING DIRECTORY FIRST
3. MAKE SURE YOU KNOW WHAT YOUR VARIABLES ARE (X and Y, ORDER MATTERS)

0. Clear your workspace.
```{r}
rm(list = ls())
# This removes everything in your environment
```

### 1. Set your working directory and load the congressional elections data (`vote_data.RData`).
```{r}
#setwd()
load("vote_data.RData")
```

Take a look at what's inside. 

Unit of analysis: incumbent-election, i.e. each row is an congressional incumbent's performance in an election.

Variable Name  | Description 
-------------  | ------------------------------------------------------------------ 
year           | year of the election
cong           | session of Congress prior to the election
idno           | unique ICPSR identifier for each incumbent
stcd           | unique idenitifier indicating the state and the congressional district
ov             | the share of two-party votes for the incumbent (if the district was contested by a Democrat and a Republican).  It is on a scale of 0-100.
ovp            | the share of two-party votes for the incumbent in the last election (if the district was contested by a Democrat and a Republican).  It is on a scale of 0-100.
opres          | two party share of presidential votes for the incumbent's party.  In the concurrent presidential election if presidential year; two years previous if a midterm. It is on a scale of 0-100.
ownexp         | campaign spending by the incumbent, in nominal dollars
oppexp         | campaign spending by the challenger, in nominal dollars
state          | state name
frosh          | freshman incumbent, facing first reelection effort
quality        | one of the candidates is an incumbent and is facing a challenger with previous experience in an elective office
incwin2        | incumbent victory dummy
stcd_redist    | stcd coupled with redistricting indicator. The last digit is 1 if the district was redistricted since the last election; 0 otherwise.
moddwnom1      | Incumbent's voting record (based on dw-nominate).  1 means extreme in the direction of the rest of his party (e.g. an extreme liberal for a Dem), -1 means extreme opposite the direction of his party (e.g.  an extreme liberal for a Rep).


### 2. How many **unique** incumbents/candidates ("idno") are in the data-set?

First have a look at your idno variable. Print it out to the console (do NOT knit it):
```{r}
#vote_data$idno  #delete the #sign to run the code
```
You can immediately tell that the ID numbers are repeated sometimes. We don't care about the duplicates, so we can use the `unique()` function which returns the non-repeating elements in a vector. The `length()` command tells you how many elements are in vector. Thus, we can combine the two functions to get the **number of unique candidates** in the data-set. 

```{r}
length(unique(vote_data$idno))
# Hint: unpack this command and print it out to the console
```


### 3. What is the average number of incumbents in the dataset across the years?
```{r}
# Average number of incumbents across the years (average times each year showing up in this dataset)
mean(table(vote_data$year))
```

Let's unpack this code. The `table()` provides a count of how many times each year shows up in the dataset. Because each observation here is an incumbent's election, this essentially creates a table of the number of incumbents for each year.

```{r}
table(vote_data$year)
```

We can then average the number of incumbents across the years by applying the `mean()` function to that.
```{r}
mean(table(vote_data$year))
```

### Another example using `table()`.

The variable incwin2 is a dummy variable for whether the incumbent won or lost.

```{r}
table(vote_data$incwin2)
```
The `prop.table()` command will give you proportions. Hint: remember that you can type `?prop.table()` in your console to get further information about R commands.

```{r}
prop.table(table(vote_data$incwin2))
```

### 4. Did incumbents outspend their opponents or not? 

Let us create a new binary variable that is equal to 1 if the opponent outspent an incumbent and 0 otherwise.

```{r}
vote_data$outspent <- ifelse(vote_data$oppexp > vote_data$ownexp, 1, 0)
```

Let us unpack this code. The syntax of the `ifelse()` function is `ifelse(condition, if condition is TRUE, if condition is FALSE)`.

Recall that `oppexp` is the expenditure by the opponent, and that `ownexp` is the expenditure by the incumbent.

The condition is always a logical statement. Let us print it to our console and see what happens (again, do NOT knit it):
```{r}
#vote_data$oppexp > vote_data$ownexp
```

Therefore, if the *condition evaluates to TRUE*, that is, if the opponent outspent the incumbent, our dummy variable `outspent` will be equal to 1. Otherwise, it will be equal to 0.

(In problem set 5 you will use a similar command, but your variable is a factor. Check leture slides for how to handle factors).
Hint: you can access more information about **operator/logical operator** via this link https://www.statmethods.net/management/operators.html.

### 5. Do most opponents outspend incumbents?
```{r}
table(vote_data$outspent)
```
We can see that most opponents DON'T outspend incumbents (0 = 4920, and 1 = 484).

What percentage of incumbents are outspent by their opponents?
```{r}
prop.table(table(vote_data$outspent))
```


### 6. Run a regression to figure out if outspending results in a higher vote share.
```{r}
model1 <- lm(ov ~ outspent, data = vote_data)
summary(model1)
# ov: the share of two party votes for the incumbent
```


### 7. Do the same but add some controls.
```{r}
model2 <- lm(ov ~ outspent + moddwnom1 + frosh, data = vote_data)
summary(model2)
#moddwnom1: Incumbent's voting record (based on dw-nominate).  1 means extreme in the direction of the rest of his party (e.g. an extreme liberal for a Dem), -1 means extreme opposite the direction of his party (e.g.  an extreme liberal for a Rep).
# frosh: freshman incumbent, facing first reelection effort
```

### 8. What is the predicted DV for a hypothetical freshman incumbent whose outspent their opponents, and "moddwnom1" level of 1?

```{r}
#One method: plug in numbers for each IV to get DV
65.0411+(-9.2151)*1 +4.5491*1+ -2.9451*1  #57.43
```
```{r}
#Another method: use predict()
predict(model2, newdata = data.frame(outspent=1, frosh=1,moddwnom1=1))
#same result:57.43004 

```

### Bonus: Add a list of predicted DVs to your dataset and make a plot with the predicted DV values
```{r}
#There are a couple of ways of doing this. 
vote_data$fit<-predict(model2)
vote_data$fit2<-fitted.values(model2)
plot(vote_data$fit~vote_data$outspent,xlab="Whether incumbent outspent opponents",ylab="Predicted voteshare")
```

